---
title: "Assignment B-1"
output:
  md_document:
    variant: markdown_github
---

In the following assignment, I created and described a function inspired by my 
personal research. I am an ecological entomologist, interested in studying how 
mosquitoes will react to global warming and climate change. 

One trait we can study in mosquitoes is body size, so I made a function which 
can quickly create a boxplot and summary statistics of the body size of insects 
(or any measured numeric variable) from different experimental growth treatments. 

The function I made below is not meant to provide publishable figures and does 
not use any statistical analyses to determine significant relationships. It is 
meant to be used at the beginning of data analysis, to gain a general 
understanding of how the data looks.

First we will begin by loading packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load packages}
library(tidyverse)
library(testthat)
```


Below is the function.

``` {r, Function}

#' Quick boxplot and summary statistics 
#'
#' 'box_and_stats()' returns a boxplot and summary statistics (max, min, mean, 
#'    median, standard deviation) for a dataset which contains categorical 
#'    treatment groups and a measured numeric variable. This function was made 
#'    to assess the body size of insects (or any measured numeric variable) from 
#'    different experimental growth treatments. 
#'
#' @param data a tidy data set. The data must be 'tidy' for this function to 
#'    work, so each column must be a variable, each row must be a unique 
#'    observation, and each cell must contain a single value.
#' @param x a column which contains different categorical treatments, and can be 
#'    characters or factors. These are the treatments groups the measured 
#'    variable will be divided into for analyses. The presence of NA values is
#'    not permitted.
#' @param y a column which contains the measured numeric variable. This is 
#'    dependent variable the summary statistics will be calculated from. The 
#'    presence of NA values is permitted, however they will be removed for analyses. 
#' @return This function will return two items. The first item being a boxplot 
#'    displaying the data from the numeric measured variable grouped by the 
#'    identified treatments. The second item is a tibble containing 6 columns: 
#'    1 column containing the different treatments, and 5 columns for the 
#'    minimum, maximum, mean, median, and standard deviation for the measured 
#'    variable from each treatment. 

box_and_stats <- function(data, x, y) {
  
   x_data_check <- dplyr::summarise(data,
                                   is_character_x = is.character({{x}}) | is.factor({{x}}),
                                   class_x = class({{x}}))
  if(!x_data_check$is_character_x) {
    stop("Selected x-column is not character or factor, column is:", x_data_check$class_x)}
   
   y_data_check <- dplyr::summarise(data,
                                   is_numeric_y = is.numeric({{y}}),
                                   class_y = class({{y}}))
  if(!y_data_check$is_numeric_y) {
    stop("Selected y-column is not numeric, column is:", y_data_check$class_y)}
   
  data_to_analyze <- data %>%
    dplyr::mutate(fun_treatment = as.factor({{x}})) %>%
    dplyr::mutate(fun_bodysize = {{y}}) %>%
    dplyr::filter(fun_treatment != "NA") %>%
    dplyr::filter(fun_bodysize != "NA")
  
  quick_boxplot <- data_to_analyze %>%
    ggplot(mapping= aes(x=fun_treatment, y=fun_bodysize))+ 
    geom_boxplot(mapping =aes(x=fun_treatment,y=fun_bodysize),width=0.5)+
    geom_jitter(mapping =aes(x=fun_treatment, y=fun_bodysize), width = 0.1, alpha = 0.6)+
    theme_minimal()+
    ylab("Body size measure") +
    xlab("Treatment")
    
  quick_stats <- data_to_analyze %>%
    dplyr::group_by(fun_treatment) %>%
    dplyr::summarize(min = min(fun_bodysize),
                     max = max(fun_bodysize),
                     mean = mean(fun_bodysize),
                     median = median(fun_bodysize),
                     SD = sd(fun_bodysize))
  
  quick_results <- list(quick_boxplot, quick_stats)
                     
  return(quick_results)
  
}
```



The following code chunk will demonstrate this function used on the built-in R 
data set 'iris'. The categorical treatment variable here will be Species, and 
the measured numeric variable will be Sepal.Width. As you can see, function 
ignores all other variables in the data set, focusing only on Species and 
Sepal.Width, returning a boxplot and brief summary statistics. 

With this quick function, I am quickly able to see that the species 'setosa' is 
larger than the species 'versicolor'. This newly found information may prompt 
me to further investigate the differences between these two species, running 
statistical tests to determine if there are any significance differences between 
species.

```{r, iris example}
box_and_stats(iris, Species, Sepal.Width)
```

Here I used the built-in 'mtcars' data set to show that the inputs for the 
function are strict. While this data set does not contain body size measured 
from different treatments, the function is flexible and can work for any data as
long as the parameters are followed. This data set can work because it is tidy 
data, however the input for "x" was neither a character nor a factor. Even 
though "cylinders" in cars can be considered categorical, in this data set R has 
it as a character class. 

```{r, mtcars example 1, error = TRUE}
box_and_stats(mtcars, cyl, wt)
```

If we would still like to use this function, we would have to convert the cyl 
column into a character or factor, and rerun the function. Below we see that the 
function now works. While some data sets may use numbers as categorical 
variables, this added step on the users behalf ensures that this function is 
used correctly by forcing the user to consciously convert numeric categorical 
data into characters or factors. 
```{r, mtcars example 2}

mtcars_new <- mtcars %>%
  mutate_at(vars(cyl), factor)

box_and_stats(mtcars_new, cyl, wt)

```

Below I am testing my function. 
```{r, testthat}
test_that("Error message with incorrect x input type", {
  expect_error(box_and_stats(mtcars, cyl, wt))
})
test_that("Error message with incorrect y input type", {
  expect_error(box_and_stats(iris, species, "hello"))
})

test_that("Returns the correct outputs", {
  subject <- box_and_stats(iris, Species, Sepal.Width)
  expect_s3_class(subject[[1]], "ggplot")
  expect_s3_class(subject[[2]], "data.frame")
})

test_that("Returns correct output", {
  expect_type(box_and_stats(iris, Species, Sepal.Width), "list")
})

#Creating two dataframes, one without NAs and one with 1 added NA for each 
#categorical variable.
df1 <- tibble(
  x = c("m", "a", "b","m", "a", "b", "m", "a", "b", "m", "a", "b", "m", "a", "b"),
  y = c(1, 1, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6, NA, NA , NA)
)
df2 <- tibble(
  x = c("m", "a", "b","m", "a", "b", "m", "a", "b", "m", "a", "b", "m", "a", "b"),
  y = c(1, 1, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6, NA, NA , NA)
)

test_that("Works the same with and without NA values in columns",{
  expect_equal(box_and_stats(df1, x, y), box_and_stats(df2,x,y))
})

```





